---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quickstats

<!-- badges: start -->
<!-- badges: end -->

The goal of quickstats is to provide simple functionality to statistically compare two groups and output the results in a publication-ready table. 

## Installation

You can install the development version of quickstats from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jtmenchaca/quickstats")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(quickstats)
library(dplyr)
## basic example code
data(iris)
summary = iris |>
  summarize_cols_by_group(
    group_col = "Species",
    continuous_cols = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"),
    rounding_digits = 1
  )


data(mtcars)
summary = mtcars |>
  summarize_cols_by_group(
    group_col = "cyl",
    binary_or_cat_cols = c("vs", "am", "gear", "carb"),
    continuous_cols = c("mpg", "disp", "hp", "drat", "wt", "qsec")
  )

```

By default, the binary/categorical columns are compared across groups using the Fisher Exact test (if two groups) and the ANOVA test (if greater than two groups).  

```{r cars}
mtcars |>
  summarize_cols_by_group(
    group_col = "cyl",
    binary_or_cat_cols = c("gear", "carb")
  )
```

The continuous columns are compared across groups using either the Student's t-Test (if following a normal distribution) or the Wilcoxon signed-rank test (if the distribution is non-normal).  The algorithm uses the Shapiro–Wilk test to test normality.

```{r }
mtcars |> 
  filter(cyl != 6) |> 
  summarize_cols_by_group(group_col = "cyl", continuous_cols = c("mpg", "disp", "hp"))
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
